Transient Corotating Clumps Around Adolescent Low-Mass Stars From Four Years of TESS

### 2023-09-04 paper/ms.tex
update ms: we are back to CPV

%%%%%%%%
% CRS %
%%%%%%%%

@@

\received{---}
\revised{---}
\accepted{---}
\shorttitle{Four Years of Complex Rotators}

@@

% 250 word max!
\begin{abstract}
  We explore a set of pre-main-sequence stars
  with sub-solar masses, rotation periods of 0.2--2 days, optical variability amplitudes
  of a few percent, and sharp recurring light curve features that last for
  hundreds of cycles.  The sharp features are probably caused by
  eclipsing clumps of dust or gas orbiting at the corotation radius.
  We refer to this class of variable star as ``complex rotators'' (CRs).
  To improve our understanding of these
  enigmatic objects, we conducted a survey for CRs using TESS
  short-cadence data collected between 2018~July and 2022~September.
  Our search of \nstarssearched\ K and M dwarfs with $T$$<$16 and
  $d$$<$150\,pc yielded \ngoods\ high-quality CRs, most of which are
  new discoveries.  Our sample includes the brightest
  ($T$$\approx$9.5), closest ($d$$\approx$20\,pc), and oldest
  ($\approx$200\,Myr) known CRs.  One of the new discoveries, LP
  12-502, stands out for the complexity of its light curve --- with as
  many as eight flux dips per cycle --- and for the detail with which the
  evolution in each dip's amplitude and phase can be tracked over
  1{,}500 cycles.  We argue that corotating gas or dust remains the
  most viable explanation for CRs. The gas or dust is probably
  entrained by the star's magnetic field, and the sharp features could
  result from a multipolar field topology, a hypothesis supported by
  correspondences between the light curves of CRs and of B stars
  that may have multipolar magnetic fields.
\end{abstract}

@@

		\vspace{-0.65cm}
		\subfloat{
			\includegraphics[width=0.66\textwidth]{f1b.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Complex rotators (CRs)}:
		{\it Top:} Phase-folded TESS light curves for three CRs.  Each
		panel shows the average of the data accumulated over one month,
		relative to the mean stellar brightness.
		Gray circles are raw 2-minute data; black circles are binned to
		300 points per cycle.  The period in hours is printed in the
		bottom right corner.  Left-to-right, the objects are LP 12-502
		(TIC 402980664; Sector~19), TIC 94088626 (Sector 10), and TIC
		425933644 (Sector~28).
		{\it Bottom:} Cartoon explanations for the phenomenon.  The dust
		clump scenario (left) and gas prominence scenario (right) both invoke
		magnetically-entrained corotating material.
	}
	\label{fig:f1}
\end{figure*}

@@

Data from K2 \citep{2014PASP..126..398H} and TESS
\citep{2015JATIS...1a4003R} have revealed a new class of variable star
for which the root cause of variability is only beginning to become
clear: complex rotators (CRs).  These objects are identified from
their optical light curves, which show nearly periodic troughs that
are either sharp or broad; these troughs are often superposed on
quasi-sinusoidal spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CRs show up to eight dips per cycle.  Most CRs are
pre-main-sequence M dwarfs with ages of $\approx$5-150 million years
(Myr), and rotation periods of 0.2--2 days.  They are observed to
comprise $\approx$1-3\% of M dwarfs younger than 100 Myr
\citep{2016AJ....152..114R,2022AJ....163..144G}.  They generally do
not show near-infrared excesses indicative of dusty disks, but the
wavelength-dependent dip amplitudes of some CRs is consistent with
reddening by dust
\citep{2017PASJ...69L...2O,2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
The dip amplitudes and phases usually evolve gradually over tens to
hundreds of cycles, although they have occasionally been observed to
change abruptly within one cycle
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

The sharp features of CR light curves can have durations as short as
5\% of the rotation period ($P_{\rm rot}$), which is too short to be
caused by starspots rotating into and out of view.  Starspots produce
flux variations with characteristic timescales of $P_{\rm rot}$ and
$0.5\,P_{\rm rot}$.  With finely-tuned viewing geometries, starspots
can produce dip durations as short as $\approx$$0.2\,P_{\rm rot}$, but
in such cases, limb darkening causes the dip amplitudes to be smaller
than the observed amplitudes of $\sim$1\% (see
\citealt{2017AJ....153..152S}, Figures 37-41).  Thus, a
``starspot-only'' scenario can be ruled out for many CRs
\citep{2017AJ....153..152S,2019ApJ...876..127Z,2021MNRAS.500.1366K}.
Given that many CRs cannot be explained by starspots alone, and
working under the assumption that all CRs share the same basic
physical scenario, we discard the ``starspot-only'' scenario.
Instead, the correct explanation probably involves material above the
stellar surface that is spatially confined with sharp boundaries
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.

@@

In both models, the corotation radius is the location at which matter
concentrates.  The empirical basis for this is that
the sharp CR features are superposed over smooth, quasi-sinusoidal
starspot profiles.  The theoretical importance of the corotation
radius has been discussed in studies of magnetic rotators
\citep[e.g.][]{1973ApJ...184..271L,1985Ap&SS.116..285N,2005ApJ...634.1214L}.
%
%including neutron stars \citep[e.g.]{1973ApJ...184..271L}, magnetic
%massive stars \citep[e.g.][]{1985Ap&SS.116..285N}, and accreting T
%Tauri stars \citep[e.g.][]{2005ApJ...634.1214L}.  
%
In regions where the magnetic field dominates the flow (i.e.~$B^2/8\pi
> \rho v^2 /2$), matter is dragged along with the field lines.  The
tendency for charged gas or dust to become trapped at corotation is
then tied to how of the four relevant forces -- gravity, Lorentz,
inertial Coriolis, and inertial centrifugal -- the Lorentz and
Coriolis only act perpendicular to field lines, while gravity and the
centrifugal force are in balance at $R_{\rm c}$ \citep[e.g.][their
Section~2]{2005MNRAS.357..251T}.  Another way to phrase this statement
is that there is a local minimum in the effective potential of the
corotating frame; given a flow from either the star or from a
tenuous accretion disk, this local potential minimum enables material
to build up \citep{2005MNRAS.357..251T}.

Although theoretical heritage for understanding rapid magnetic
rotators exists, CRs have remained mysterious because they have been both
hard to discover and hard to characterize.   They are hard to discover
because they are rare: CRs comprise $\approx$1\% of the youngest
$\approx$1\% of M dwarfs \citep{2018AJ....155..196R}.  Out of the
millions of stars monitored by K2 and TESS, about 50 CRs have been
reported to date
\citep{2016AJ....152..114R,2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z,2020AJ....160...86B,2021AJ....161...60S,2022AJ....163..144G,2023ApJ...945..114P}.
They have been hard to characterize because many of the known CRs are
faint; the initial K2 discoveries
\citep{2016AJ....152..114R,2017AJ....153..152S} were M2-M6 dwarfs at
distances $\gtrsim$100\,pc, yielding optical brightnesses of
$V$$\approx$15.5 to $V$$>$20.  At such magnitudes, high-resolution
time-series spectroscopy is out of reach with current facilities,
despite the potential utility of such observations.

In this work, we aim to find bright and nearby CRs, since these
objects will be the most amenable to detailed photometric and
spectroscopic analyses.  To do this, we use 120-second cadence data
acquired by TESS between 2018 July and 2022 September (Sectors 1-55;
Cycles 1-4).  We present our search methods in
Section~\ref{sec:methods}, and the resulting CR catalog in
Section~\ref{sec:results}.  The observed evolution of many CRs over a
two-year baseline is described in Section~\ref{sec:evoln}, including a
deep-dive into the behavior of an especially interesting object, LP
12-502.  We discuss a few implications in
Section~\ref{sec:discussion}, and conclude in
Section~\ref{sec:conclusion}.

Some comments on nomenclature are needed.  What we are calling
``complex rotators''
\citep{2019ApJ...876..127Z,2022AJ....163..144G,2023ApJ...945..114P}
have also been called ``transient flux dips'', ``persistent flux
dips'', and ``scallop shells'' by \citet{2017AJ....153..152S}, or more
generally ``complex periodic variables'' by
\citet{2023MNRAS.518.2921K}.  The CRs should not be conflated with
``dippers'', which are classical T Tauri stars with infrared excesses,
and which show large-amplitude variability linked to obscuring inner
disk structures and accretion hot spots
\citep{2014AJ....147...82C,2021ApJ...908...16R}.  The phenomenology
and stellar properties of CRs and dippers are quite different (though
see Sections~\ref{subsec:irexcess} and~\ref{subsec:discdippers}).
While we considered other possibilities, we ultimately concluded that
``complex rotator'' correctly draws attention to the fact that
{\it complexity} relative to quasi-sinusoidal starspots is the
defining phenomenological feature of these objects, while also highlighting the expectation that {\it
rotation} will play a central role in explaining their physical
behavior.

@@

To construct our sample, we analyzed the short-cadence data acquired
by TESS between 2018 July 25 and 2022 September 1 (Sectors 1-55).
Specifically, we used the 120-second cadence light curves produced by
the Science Processing and Operations Center at the NASA Ames Research
Center \citep{2016SPIE.9913E..3EJ}.  While the TESS data products from
these sectors also included full frame images with cadences of 600 and
1800 seconds for a larger number of sources, we restricted our
attention to the 120-second data for the sake of uniformity and
simplicity in data handling.  In exchange, we sacrificed both
completeness and homogeneity of the selection function.  While TESS
cumulatively observed $\approx$90\% of the sky for at least one lunar
month between 2018 July and 2022 September, the 120-second cadence
data were collected for only a subset of observable stars that were
preferentially nearby and bright \citep[see][]{2021PASP..133i5002F}.
The total 120-second data volume
from Sectors 1-55 included 1{,}087{,}475 short-cadence light curves, which were
available for 428{,}121 unique stars.
% latter stats from MERGED_spoc2min_sector1_to_sector58.csv

To further simplify our search, we defined our target sample as
stars with 120-second cadence TESS light curves satisfying the
following four conditions:
\begin{align}
  T &< 16 \quad&(\mathrm{Amenable\ with\ TESS}) \label{eq:one}\\
  \bprp &> 1.5 \quad&(\mathrm{Red\ stars\ only})\\
  M_{\rm G} &> 4 \quad&(\mathrm{Dwarf\ stars\ only})\\
  d &< 150\,{\rm pc} \quad&(\mathrm{Close\ stars\ only}) \label{eq:four}.
\end{align}
Here, $M_{\rm G} = G + 5\log(\varpi_{\rm as}) + 5$ is the Gaia
$G$-band absolute magnitude, $\varpi_{\rm as}$ is the parallax in
units of arcseconds, and $d$ is a geometric distance defined by
inverting the parallax and ignoring any zero-point correction.  We
performed this selection by cross-matching TIC8.2
\citep{2019AJ....158..138S,2021arXiv210804778P} against the Gaia DR2
point-source catalog \citep{2018A&A...616A...1G}.  We opted for Gaia
DR2 rather than DR3 because the base catalog for TIC8 was Gaia DR2,
which facilitated a one-to-one crossmatch using the Gaia source
identifiers.  The target sample ultimately included \nstarssearched\ M
dwarfs and late-K dwarfs, down to $T$$<$16 and out to $d$$<$150\,pc.
For stars with multiple sectors of TESS data available, we searched
for CR signals independently.
In total, our \nstarssearched\ star target list included \nlcssearched\ month-long light
curves.

@@

\subsection{CR discovery}
\label{subsec:discoverymethods}

Prior to this study, most CRs have been found by visually examining
all the light curves of stars in young clusters
\citep{2016AJ....152..114R,2017AJ....153..152S,2023ApJ...945..114P},
or by flagging light curves with short periods and a large number of
strong Fourier harmonics for visual inspection
\citep{2019ApJ...876..127Z}.  In this work, we implemented a new
search approach based on counting the number of sharp local minima in
phase-folded light curves, while also using the Fourier approach.  We
applied these two search techniques independently.

@@

The dip counting technique aims to count sharp local minima in
phase-folded light curves.  The most remarkable CRs often show three
or more dips per cycle, which distinguishes them from other types of
variables such as synchronized and spotted binaries (RS CVn stars).

@@

Wrapping is necessary to eliminate edge effects when fitting the light
curve and when identifying local minima in the residuals.  A phased
light curve would usually cover phases $\phi \in [ 0,1 ]$.  We instead
performed the analysis described above using a phase-folded light
curve spanning $\phi \in [-1,2 ]$, which was created by duplicating
and concatenating the ordinary phase-folded light curve.  The free
parameters we adopted throughout the analysis -- for instance the
maximum number of spline knots per cycle, and the height and depth
criteria for dips -- were chosen during testing based on the desire to
correctly re-identify a large fraction ($>$90\%) of previously known
CRs, while also being able to consistently reject common false
positives such as spot-induced variability and
eclipsing binaries.

In short, CR candidates were identified by requiring a peak PDM period
below two days and the presence of at least three sharp local minima,
based on at least one sector of the TESS 120-second data. Candidates
were then inspected visually as described in
Section~\ref{subsec:visual}.

@@

If a peak of such significance was found, we generated a ``summary
sheet'' with information about the star, its full and folded light
curves, Fourier transform, potential contaminating stars, and
information about these contaminating stars.  Summary sheets were
generated for $\approx$10\% of the 20{,}000 120-second targets in each
sector.  We reviewed each summary sheet and tried to classify each
light curve based on visual inspection of its morphology (with
categories including eclipsing binary, CR, RS CVn, and cataclysmic
variable). Whenever a CR was identified, the light curves from
multiple sectors were stitched together and analyzed jointly, for a
more accurate period determination.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{f2.pdf}
		\vspace{-0.3cm}
		\caption{
      {\bf Validation plots used to classify CRs}.  The complete figure
      set, with one image per sector for each of the \nallcands\
      objects in Table~\ref{tab:thetable} is available in the online
      journal. {\bf For internal collaboration review:
      \url{https://www.dropbox.com/sh/khtwk5a6z0zgrif/AABRG21Ire7VS4BUnaD1hOR8a?dl=0}}.
      Panels are as follows.
			{\it a)}: Phase-folded light curve; gray points are raw 2-minute
			data and black points are binned to 200 points per cycle.
      The adopted period is given in the lower-right corner.
			{\it b)}: Phase-dispersion minimization (PDM) periodogram.
			Dotted lines show up to the 10$^{\rm th}$ harmonic and
			subharmonic.
      {\it c)}: DSS finder chart, with 21$''$ and 42$''$ radius
      circles for scale.  One TESS pixel has a full side length of
      21$''$.
			{\it d)}: Cleaned light curve, binned to 20-minute cadence, in
			Barycentric TESS Julian Date (BTJD).
			{\it e)}: Phase-folded light curve, binned to 100 points per
			cycle.  The gray line denotes the spline-fit to the
			wrapped phase-folded light curve, and small gray triangles
			denote automatically identified local minima.
			{\it f)}: Phase-folded light curve at twice the peak period.
			{\it g)}: Phase-folded light curve at half the peak period.
			{\it h)}: Phase-folded time-series within the ``background''
			aperture defined in the SPOC light curves.
			{\it i)}: Phase-folded flux-weighted centroid in the column
			direction.
			{\it j)}: Phase-folded flux-weighted centroid in the row
			direction.
      {\it k)}: Gaia DR2 color--absolute magnitude diagram. The gray
      background denotes stars within 100\,pc from
      \citet{2021A&A...649A...6G}.
      {\it l)}: Information from Gaia DR2, TIC8, and the automated
      dip-counting search pipeline.  ``Neighbors'', abbreviated
      ``nbhr'', are listed within apparent distances of 2 TESS pixels
      if $\Delta T$$<$2.5.
      {\it m)}: BANYAN-$\Sigma$ v1.2 association probabilities,
      calculated using positions, proper motions, and the parallax.
		}
		\label{fig:vet}
	\end{center}
\end{figure*}

We assessed whether the objects found using the dip-counting
(Section~\ref{subsec:counting}) and Fourier
(Section~\ref{subsec:fourier}) techniques were consistent with
expectations for CRs by assembling the data shown in
Figure~\ref{fig:vet}.  We labeled a star as a ``good'' CR if it met
all of the following criteria for at least one TESS sector:
\vspace{-2pt}
\begin{itemize}[leftmargin=*]
  \setlength\itemsep{-2pt}
  \item $P<2$\,days.
  %
  \item At least three dips per cycle, or else otherwise
    oddly-shaped dips.
  %
  \item Dip phases (mostly) repeat over $\approx$30\,day sector duration.
\end{itemize}
\vspace{-2pt}
We also noted ''possible'' CRs for which the classification was more
ambiguous, and ``false positives'' that are definitely not CRs.  The
most common false positives for both the Fourier and dip-counting
techniques were eclipsing binaries, ordinary spotted rapid rotators,
and complex light curves due to multiple stars contributing to the
photometric aperture.  Our dip-counting pipeline flagged
\nuniqdipflagged\ unique stars for visual inspection; about 20\% were
subsequently labeled either good or possible CRs.  From the Fourier
pipeline, $\approx$0.5\% of stars that passed the 12-$\sigma$ peak
threshold were eventually classified as CRs.  This is because the
Fourier pipeline was implemented as a general variability search; its
results are also being used for analyses on topics other than CRs.

@@

\subsection{CR catalog}
\label{sec:catalog}

>>>>

%%%%%%%%
% CPVS %
%%%%%%%%

@@

\received{---}
\revised{---}
\accepted{---}
\shorttitle{Four Years of Complex Periodic Variables}

@@

% 250 word max!
\begin{abstract}
  We explore a set of pre-main-sequence stars
  with sub-solar masses, rotation periods of 0.2--2 days, optical variability amplitudes
  of a few percent, and sharp recurring light curve features that last for
  hundreds of cycles.  The sharp features are probably caused by
  eclipsing clumps of dust or gas orbiting at the corotation radius.
  We refer to this class of star as ``complex periodic variables'' (CPVs).
  To improve our understanding of these
  enigmatic objects, we conducted a survey for CPVs using TESS
  short-cadence data collected between 2018~July and 2022~September.
  Our search of \nstarssearched\ K and M dwarfs with $T$$<$16 and
  $d$$<$150\,pc yielded \ngoods\ high-quality CPVs, most of which are
  new discoveries.  Our sample includes the brightest
  ($T$$\approx$9.5), closest ($d$$\approx$20\,pc), and oldest
  ($\approx$200\,Myr) known CPVs.  One of the new discoveries, LP
  12-502, stands out for the complexity of its light curve --- with as
  many as eight flux dips per cycle --- and for the detail with which the
  evolution in each dip's amplitude and phase can be tracked over
  1{,}500 cycles.  We argue that corotating gas or dust remains the
  most viable explanation for CPVs. The gas or dust is probably
  entrained by the star's magnetic field, and the sharp features could
  result from a multipolar field topology, a hypothesis supported by
  correspondences between the light curves of CPVs and of B stars
  that may have multipolar magnetic fields.
\end{abstract}

@@

		\vspace{-0.65cm}
		\subfloat{
			\includegraphics[width=0.66\textwidth]{f1b.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Complex periodic variables (CPVs)}:
		{\it Top:} Phase-folded TESS light curves for three CPVs.  Each
		panel shows the average of the data accumulated over one month,
		relative to the mean stellar brightness.
		Gray circles are raw 2-minute data; black circles are binned to
		300 points per cycle.  The period in hours is printed in the
		bottom right corner.  Left-to-right, the objects are LP 12-502
		(TIC 402980664; Sector~19), TIC 94088626 (Sector 10), and TIC
		425933644 (Sector~28).
		{\it Bottom:} Cartoon explanations for the phenomenon.  The dust
		clump scenario (left) and gas prominence scenario (right) both invoke
		magnetically-entrained corotating material.
	}
	\label{fig:f1}
\end{figure*}

@@

Data from K2 \citep{2014PASP..126..398H} and TESS
\citep{2015JATIS...1a4003R} have revealed a new class of variable star
for which the root cause of variability is only beginning to become
clear: complex periodic variables (CPVs).  These objects are identified from
their optical light curves, which show nearly periodic troughs that
are either sharp or broad; these troughs are often superposed on
quasi-sinusoidal spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CPVs show up to eight dips per cycle.  Most CPVs are
pre-main-sequence M dwarfs with ages of $\approx$5-150 million years
(Myr), and rotation periods of 0.2--2 days.  They are observed to
comprise $\approx$1-3\% of M dwarfs younger than 100 Myr
\citep{2016AJ....152..114R,2022AJ....163..144G}.  They generally do
not show near-infrared excesses indicative of dusty disks, but the
wavelength-dependent dip amplitudes of some CPVs is consistent with
reddening by dust
\citep{2017PASJ...69L...2O,2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
The dip amplitudes and phases usually evolve gradually over tens to
hundreds of cycles, although they have occasionally been observed to
change abruptly within one cycle
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

The sharp features of CPV light curves can have durations as short as
5\% of the rotation period ($P_{\rm rot}$), which is too short to be
caused by starspots rotating into and out of view.  Starspots produce
flux variations with characteristic timescales of $P_{\rm rot}$ and
$0.5\,P_{\rm rot}$.  With finely-tuned viewing geometries, starspots
can produce dip durations as short as $\approx$$0.2\,P_{\rm rot}$, but
in such cases, limb darkening causes the dip amplitudes to be smaller
than the observed amplitudes of $\sim$1\% (see
\citealt{2017AJ....153..152S}, Figures 37-41).  Thus, a
``starspot-only'' scenario can be ruled out for many CPVs
\citep{2017AJ....153..152S,2019ApJ...876..127Z,2021MNRAS.500.1366K}.
Given that many CPVs cannot be explained by starspots alone, and
working under the assumption that all CPVs share the same basic
physical scenario, we discard the ``starspot-only'' scenario.
Instead, the correct explanation probably involves material above the
stellar surface that is spatially confined with sharp boundaries
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.

@@

In both models, the corotation radius is the location at which matter
concentrates.  The empirical basis for this is that
the sharp CPV features are superposed over smooth, quasi-sinusoidal
starspot profiles.  The theoretical importance of the corotation
radius has been discussed in studies of magnetic rotators
\citep[e.g.][]{1973ApJ...184..271L,1985Ap&SS.116..285N,2005ApJ...634.1214L}.
%
%including neutron stars \citep[e.g.]{1973ApJ...184..271L}, magnetic
%massive stars \citep[e.g.][]{1985Ap&SS.116..285N}, and accreting T
%Tauri stars \citep[e.g.][]{2005ApJ...634.1214L}.  
%
In regions where the magnetic field dominates the flow (i.e.~$B^2/8\pi
> \rho v^2 /2$), matter is dragged along with the field lines.  The
tendency for charged gas or dust to become trapped at corotation is
then tied to how of the four relevant forces -- gravity, Lorentz,
inertial Coriolis, and inertial centrifugal -- the Lorentz and
Coriolis only act perpendicular to field lines, while gravity and the
centrifugal force are in balance at $R_{\rm c}$ \citep[e.g.][their
Section~2]{2005MNRAS.357..251T}.  Another way to phrase this statement
is that there is a local minimum in the effective potential of the
corotating frame; given a flow from either the star or from a
tenuous accretion disk, this local potential minimum enables material
to build up \citep{2005MNRAS.357..251T}.

Although theoretical heritage for understanding rapid magnetic
rotators exists, CPVs have remained mysterious because they have been both
hard to discover and hard to characterize.   They have been hard to discover
because they are rare: CPVs comprise $\approx$1\% of the youngest
$\approx$1\% of M dwarfs \citep{2018AJ....155..196R}.  Out of the
millions of stars monitored by K2 and TESS, about 50 CPVs have been
reported to date
\citep{2016AJ....152..114R,2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z,2020AJ....160...86B,2021AJ....161...60S,2022AJ....163..144G,2023ApJ...945..114P}.
They have been hard to characterize because many of the known CPVs are
faint; the initial K2 discoveries
\citep{2016AJ....152..114R,2017AJ....153..152S} were M2-M6 dwarfs at
distances $\gtrsim$100\,pc, with optical brightnesses of
$V$$\approx$15.5 to $V$$>$20.  At such magnitudes, high-resolution
time-series spectroscopy is out of reach with current facilities,
despite the potential utility of such observations.

In this work, we aim to find bright and nearby CPVs, since these
objects will be the most amenable to detailed photometric and
spectroscopic analyses.  To do this, we use 120-second cadence data
acquired by TESS between 2018 July and 2022 September (Sectors 1-55;
Cycles 1-4).  We present our search methods in
Section~\ref{sec:methods}, and the resulting CPV catalog in
Section~\ref{sec:results}.  The observed evolution of many CPVs over a
two-year baseline is described in Section~\ref{sec:evoln}, including a
deep-dive into the behavior of an especially interesting object, LP
12-502.  We discuss a few implications in
Section~\ref{sec:discussion}, and conclude in
Section~\ref{sec:conclusion}.

Some comments on nomenclature are needed.  What we are calling
``complex periodic variables'' \citep{2023MNRAS.518.2921K} have also
been called ``complex rotators''
\citep{2019ApJ...876..127Z,2022AJ....163..144G,2023ApJ...945..114P},
``transient flux dips'', ``persistent flux dips'', and ``scallop
shells'' \citep{2017AJ....153..152S}.  The CPVs should not be conflated
with ``dippers'', which are classical T Tauri stars with infrared
excesses, and which show large-amplitude variability linked to
obscuring inner disk structures and accretion hot spots
\citep{2014AJ....147...82C,2021ApJ...908...16R}.  The phenomenology
and stellar properties of CPVs and dippers are quite different (though
see Sections~\ref{subsec:irexcess} and~\ref{subsec:discdippers}).  The
defining phenomenological features of the CPVs are that their
light curves are {\it complex}, relative to quasi-sinusoidal
starspots, and the complex features are {\it periodic}, over
timescales of order one hundred cycles.  While rotation likely
does play a central role in explaining their physical behavior, the
acronym for ``complex rotator'' is already widely used in the astrophysical
literature for cosmic rays.  Given these
considerations, we refer to the stars as complex periodic variables
(CPVs); our preferred explanation for their behavior is that transient
clumps of gas or dust orbit at the corotation radius.

@@

We searched for CPVs by analyzing the short-cadence data acquired
by TESS between 2018 July 25 and 2022 September 1 (Sectors 1-55).
Specifically, we used the 120-second cadence light curves produced by
the Science Processing and Operations Center at the NASA Ames Research
Center \citep{2016SPIE.9913E..3EJ}.  While the TESS data products from
these sectors also included full frame images with cadences of 600 and
1800 seconds for a larger number of sources, we restricted our
attention to the 120-second data for the sake of uniformity and
simplicity in data handling.  In exchange, we sacrificed both
completeness and homogeneity of the selection function.  While TESS
cumulatively observed $\approx$90\% of the sky for at least one lunar
month between 2018 July and 2022 September, the 120-second cadence
data were collected for only a subset of observable stars that were
preferentially nearby and bright \citep[see][]{2021PASP..133i5002F}.
The total 120-second data volume
from Sectors 1-55 included 1{,}087{,}475 short-cadence light curves, which were
available for 428{,}121 unique stars.
% latter stats from MERGED_spoc2min_sector1_to_sector58.csv

To simplify our search, we defined our target sample as
stars with 120-second cadence TESS light curves satisfying the
following four conditions:
\begin{align}
  T &< 16 \quad&(\mathrm{Amenable\ with\ TESS}) \label{eq:one}\\
  \bprp &> 1.5 \quad&(\mathrm{Red\ stars\ only})\\
  M_{\rm G} &> 4 \quad&(\mathrm{Dwarf\ stars\ only})\\
  d &< 150\,{\rm pc} \quad&(\mathrm{Close\ stars\ only}) \label{eq:four}.
\end{align}
Here, $M_{\rm G} = G + 5\log(\varpi_{\rm as}) + 5$ is the Gaia
$G$-band absolute magnitude, $\varpi_{\rm as}$ is the parallax in
units of arcseconds, and $d$ is a geometric distance defined by
inverting the parallax and ignoring any zero-point correction.  We
performed this selection by cross-matching TIC8.2
\citep{2019AJ....158..138S,2021arXiv210804778P} against the Gaia DR2
point-source catalog \citep{2018A&A...616A...1G}.  We opted for Gaia
DR2 rather than DR3 because the base catalog for TIC8 was Gaia DR2,
which facilitated a one-to-one crossmatch using the Gaia source
identifiers.  The target sample ultimately included \nstarssearched\ M
dwarfs and late-K dwarfs, down to $T$$<$16 and out to $d$$<$150\,pc.
For stars with multiple sectors of TESS data available, we searched
for CPV signals independently.
In total, our \nstarssearched\ star target list included \nlcssearched\ month-long light
curves.

@@

\subsection{CPV discovery}
\label{subsec:discoverymethods}

Prior to this study, most CPVs have been found by visually examining
all the light curves of stars in young clusters
\citep{2016AJ....152..114R,2017AJ....153..152S,2023ApJ...945..114P},
or by flagging light curves with short periods and a large number of
strong Fourier harmonics for visual inspection
\citep{2019ApJ...876..127Z}.  In this work, we implemented a new
search approach based on counting the number of sharp local minima in
phase-folded light curves, while also using the Fourier approach.  We
applied these two search techniques independently.

@@

The dip counting technique aims to count sharp local minima in
phase-folded light curves.  The most remarkable CPVs often show three
or more dips per cycle, which distinguishes them from other types of
variables such as synchronized and spotted binaries (RS CVn stars).

@@

Wrapping is necessary to eliminate edge effects when fitting the light
curve and when identifying local minima in the residuals.  A phased
light curve would usually cover phases $\phi \in [ 0,1 ]$.  We instead
performed the analysis described above using a phase-folded light
curve spanning $\phi \in [-1,2 ]$, which was created by duplicating
and concatenating the ordinary phase-folded light curve.  The free
parameters we adopted throughout the analysis -- for instance the
maximum number of spline knots per cycle, and the height and depth
criteria for dips -- were chosen during testing based on the desire to
correctly re-identify a large fraction ($>$90\%) of previously known
CPVs, while also being able to consistently reject common false
positives such as spot-induced variability and
eclipsing binaries.

In short, CPV candidates were identified by requiring a peak PDM period
below two days and the presence of at least three sharp local minima,
based on at least one sector of the TESS 120-second data. Candidates
were then inspected visually as described in
Section~\ref{subsec:visual}.

@@

If a peak of such significance was found, we generated a ``summary
sheet'' with information about the star, its full and folded light
curves, Fourier transform, potential contaminating stars, and
information about these contaminating stars.  Summary sheets were
generated for $\approx$10\% of the 20{,}000 120-second targets in each
sector.  We reviewed each summary sheet and tried to classify each
light curve based on visual inspection of its morphology (with
categories including eclipsing binary, CPV, RS CVn, and cataclysmic
variable). Whenever a CPV was identified, the light curves from
multiple sectors were stitched together and analyzed jointly, for a
more accurate period determination.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{f2.pdf}
		\vspace{-0.3cm}
		\caption{
      {\bf Validation plots used to classify CPVs}.  The complete figure
      set, with one image per sector for each of the \nallcands\
      objects in Table~\ref{tab:thetable} is available in the online
      journal. {\bf For internal collaboration review:
      \url{https://www.dropbox.com/sh/khtwk5a6z0zgrif/AABRG21Ire7VS4BUnaD1hOR8a?dl=0}}.
      Panels are as follows.
			{\it a)}: Phase-folded light curve; gray points are raw 2-minute
			data and black points are binned to 200 points per cycle.
      The adopted period is given in the lower-right corner.
			{\it b)}: Phase-dispersion minimization (PDM) periodogram.
			Dotted lines show up to the 10$^{\rm th}$ harmonic and
			subharmonic.
      {\it c)}: DSS finder chart, with 21$''$ and 42$''$ radius
      circles for scale.  One TESS pixel has a full side length of
      21$''$.
			{\it d)}: Cleaned light curve, binned to 20-minute cadence, in
			Barycentric TESS Julian Date (BTJD).
			{\it e)}: Phase-folded light curve, binned to 100 points per
			cycle.  The gray line denotes the spline-fit to the
			wrapped phase-folded light curve, and small gray triangles
			denote automatically identified local minima.
			{\it f)}: Phase-folded light curve at twice the peak period.
			{\it g)}: Phase-folded light curve at half the peak period.
			{\it h)}: Phase-folded time-series within the ``background''
			aperture defined in the SPOC light curves.
			{\it i)}: Phase-folded flux-weighted centroid in the column
			direction.
			{\it j)}: Phase-folded flux-weighted centroid in the row
			direction.
      {\it k)}: Gaia DR2 color--absolute magnitude diagram. The gray
      background denotes stars within 100\,pc from
      \citet{2021A&A...649A...6G}.
      {\it l)}: Information from Gaia DR2, TIC8, and the automated
      dip-counting search pipeline.  ``Neighbors'', abbreviated
      ``nbhr'', are listed within apparent distances of 2 TESS pixels
      if $\Delta T$$<$2.5.
      {\it m)}: BANYAN-$\Sigma$ v1.2 association probabilities,
      calculated using positions, proper motions, and the parallax.
		}
		\label{fig:vet}
	\end{center}
\end{figure*}

We assessed whether the objects found using the dip-counting
(Section~\ref{subsec:counting}) and Fourier
(Section~\ref{subsec:fourier}) techniques were consistent with
expectations for CPVs by assembling the data shown in
Figure~\ref{fig:vet}.  We labeled a star as a ``good'' CPV if it met
all of the following criteria for at least one TESS sector:
\vspace{-2pt}
\begin{itemize}[leftmargin=*]
  \setlength\itemsep{-2pt}
  \item $P<2$\,days.
  %
  \item At least three dips per cycle, or else otherwise
    oddly-shaped dips.
  %
  \item Dip phases mostly repeat over $\approx$30\,day sector duration.
\end{itemize}
\vspace{-2pt}
We also noted ''possible'' CPVs for which the classification was more
ambiguous, and ``false positives'' that are definitely not CPVs.  The
most common false positives for both the Fourier and dip-counting
techniques were eclipsing binaries, ordinary spotted rapid rotators,
and complex light curves due to multiple stars contributing to the
photometric aperture.  Our dip-counting pipeline flagged
\nuniqdipflagged\ unique stars for visual inspection; about 20\% were
subsequently labeled either good or possible CPVs.  From the Fourier
pipeline, $\approx$0.5\% of stars that passed the 12-$\sigma$ peak
threshold were eventually classified as CPVs.  This is because the
Fourier pipeline was implemented as a general variability search; its
results are also being used for analyses on topics other than CPVs.

@@

\subsection{CPV catalog}
\label{sec:catalog}
